\documentclass{article}
\usepackage{hyperref}
\usepackage[margin=1in]{geometry}
\usepackage{indentfirst}   % Indents first paragraph. change if u want ig
\usepackage{setspace}
\usepackage{qtree}
\doublespacing

\begin{document}
\title{\textbf{Unrolling a Recurrence}}
\author{Remington Greko, Tyler Gutowski, and Spencer Hirsch}
\date{\today}

\maketitle

\noindent Work with your team to write a report showing your knowledge of
the unrolling technique to solve a recurrence.

\smallskip

\noindent Submit the team's report of Canvas. Include a matrix indicating
who did what.

\noindent \textit{Sums}\\
\noindent Give a closed form expression for the sums below.

\begin{enumerate}
    \item \begin{equation}
            \sum_{i=0}^{i=n-1}1
          \end{equation}
    \item \begin{equation}
            \sum_{i=0}^{i=n-1}2^i
          \end{equation}
    \item \begin{equation}
            \sum_{k=0}^{k=n}(n!/k!(n-k)!)
          \end{equation}
    \item \begin{equation}
            \sum_{k=0}^{k=n-1}r^k
          \end{equation}
    \item \begin{equation}
            \sum_{i=0}^{i=n-1}kr^{k-1}
          \end{equation}
\end{enumerate}

\noindent \textbf{Unrolling a Recurrence}\\

\noindent I like this recurrence solving technique because it only involves
simple algebraic substitutions and an ability to compute closed form expressions
for common sums. Here is an example of unrolling that you can use as a template
for other problems.

Consider the recurrence:
\[ T(n) = T(n/2) + 1, T(0) = 1 \]
The idea is simple: if
\[ T(n) = T(n/2) + 1, T(1) = 0 \]
then by substitution (unrolling)
\[ T(n/2) = T(n/4) + 1, T(1) = 0 \]

\noindent n can be divded by 2 a only a finite number of times before 1 is
reached ending the unrolling. (It is olay to assume n = $2^p$ or p = logn and the
unrolling ends after p stages.) In another step
\[ T(n) = T(n/4) + 1 + 1, T(1) = 0 \]

\noindent Convince yourself this results in adding 1 to itself p times computing
the sume of values in the first column of Pascal's triangle.\\
Therefore,
\[ T(n) = p = log(n) \]

\noindent Make the connection that the recurrence models the binary search algorithm.\\
\noindent Here are some additional problems to flex and sharpen your skills.

\begin{enumerate}
    \item Solve the recurrence
            \[ T(n) = T(n-1) + n, T(0) = 1 \]
            \[ T(n) = T(n-1)^2 + n\]
            \[ T(n) = n^2 -2n + 1 + n\]
            \[ T(n) = n^2 - (2-1)n + 1\]
            \[ T(n) = n^2 - (2-1)n + 1\]
            \[T(n) = n^2\]

            Therefore, \[T(n) = n^2\]

            Mention the algorithm that is described by this recurrence.

            Bubble sort is an algorithm that has a recurrence of O($n^2$).
    \item Solve the recurrence
            \[ T(n) = 2T(n/2) + n, T(0) = 1 \]
            Mention the algorithm that is described by this recurrence.
    \item Solve the recurrence
            \[ T(n) = 2T(n-1) + 1, T(1) = 1 \]
            What famous problem is described by this recurrence?
\end{enumerate}


\pagebreak

\begin{center}
        \begin{tabular}{|p{3cm}|p{6cm}|}
            \hline
            \textbf{Name} & \textbf{Section} \\
            \hline
            Remington Greko & \\
            \hline
            Tyler Gutowski & \\
            \hline
            Spencer Hirsch & Unrolling a recurrence \\
            \hline
        \end{tabular}
    \end{center}

\end{document}